Assessment and optimization of respiratory syncytial virus prophylaxis regimens in Connecticut, \Sexpr{formatYear(startDate)}-\Sexpr{formatYear(endDate)}

### 2018-01-21 Thesis/Thesis.rnw
Secondary axis for epi weeks; text tweaks

% TODO: Plot regression line in onset/offset analysis
% TODO Erin: CL vs CI in abstract?
% TODO clarify all years analysis in figure captions
% TODO generally make figures stand more on their own
% TODO Put 1-2% punchline in the abstract, maybe
% TODO mention misalignment in results

@@

\abstractsection{Background}Respiratory syncytial virus (RSV) causes seasonal respiratory infection with potentially serious complications in children, leading to hospitalization rates as high as 50\% in high-risk infants. Safe and effective immunoprophylaxis (palivizumab) is available, but costly. The American Academy of Pediatrics (AAP) recommends palivizumab only for infants at high risk of complications, and only during the RSV season. Although the current AAP guidelines acknowledge the existence of spatial and temporal variation in RSV incidence, they do not recommend spatial or temporal adjustments to immunoprophylaxis regimen outside of Florida.

@@

\abstractsection{Results}Using preventable fraction to assess different immunoprophylaxis regimens, we found that (at \Sexpr{level * 100}\% confidence level), regimens adjusted for county-level variation in RSV incidence are superior to the AAP-recommended regimen. When we rounded the timing of those regimens to pragmatic calendrical boundaries (weekly, biweekly, and monthly), their gains over the AAP regimen diminished to.

\abstractpara{}Our best-performing spatially-adjusted pragmatic alternate to the AAP guidelines was the one based on regional RSV season midpoint with biweekly rounding. In comparison to the AAP recommendation, whose preventable fraction is \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}), our best alternate yields improvement to \Sexpr{regionStrategyCoverage("all", "countyMiddle", 2)} (\Sexpr{regionStrategyCoverageCI("all", "countyMiddle", 2, level)}) preventable fraction.

\abstractpara{}We also found that alternate regimens adjusted for temporal variation in incidence at the annual level have non-superior preventable fraction to our spatially-adjusted regimens.

@@

RSV seasonality has been established throughout the United States, based on National Respiratory and Enteric Virus Surveillance System data. In all states except for Florida, it shows a stable season onset between mid-November and early January and a stable season offset between mid-March and late April in all states except for Florida.\cite{CentersforDiseaseControlandPreventionCDC:2011wt,CentersforDiseaseControlandPreventionCDC:2013wv}

@@

Motivated by the AAP's recognition of the impact of spatial variation in RSV incidence on optimizing RSV immunoprophylaxis regimens, and the known spatial variability of RSV incidence in Connecticut, our study asks these questions:

@@

\begin{itemize}
    \item One region for the entire state
    \item One region for each of the counties with high RSV case counts: Hartford County, Fairfield County, and New Haven County. (Figure \ref{fig:map})
    \item One discontiguous region comprised of all remaining counties in Connecticut: \Sexpr{gsub("(.*), (.*)", "\\1, and \\2", paste(lowIncidenceCounties, collapse=", "))} (``low-RSV counties'').
\end{itemize}

@@

@@

\subsection{Evaluating spatially-adjusted prophylaxis regimens}

@@

In each variant, we again estimated the preventable fraction of all six spatially-adjusted regimens.

@@

We then compared those regimens to the one recommended by AAP and to the six spatially-adjusted alternates (with no calendrical rounding) discussed above.

@@

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=county), width=.1) + 
        geom_point(aes(y=median, x=county), size=.75) + 
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        labs(aes(x=variable)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_discrete(labels=countyLabels) + 
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks))) + 
        labs(y=NULL) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=30, hjust=1)
        )

@@

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_ribbon(aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper, x=time), fill=grey(.25)) +
    	geom_segment(aes(x=start, y=lineYmin, xend=start, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=end, y=lineYmin, xend=end, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_rect(aes(xmin=start, ymin=barYmin, xmax=end, ymax=barYmax, fill=variable, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=-Inf, y=unprotected.start, xend=ppx.start, yend=unprotected.start), linetype="11", data=thresholds, size=.375) + 
        geom_segment(aes(x=-Inf, y=unprotected.end, xend=ppx.end, yend=unprotected.end), linetype="11", data=thresholds, size=.375) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.end), data=thresholds, shape=18) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.start), data=thresholds, shape=18) + 
        scale_y_continuous(labels=latexPercent, breaks=seq(0, 1, by=.2), limits=c(-0.1, 1.1), expand=c(0, 0), minor_breaks=epiWeekBreaks) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        scale_color_manual(labels=spanLabels, values=twoToneAlt) +
        scale_fill_manual(labels=spanLabels, values=twoToneAlt) +
        guides(color=guide_legend(keywidth=3, keyheight=.4)) + 
        facet_wrap(~county, ncol=5, labeller=labeller(county=countyLabels)) +
        labs(x=NULL, y="Cumulative indicence fraction") + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
            legend.position="bottom"
        )

@@

\setcounter{figure}{4}
\begin{figure*}
\begin{center}
\input{\figuresDir/coverageByRegionRegimen.tex}
\caption[Comparison of six spatially-adjusted RSV prophylaxis regimens to the AAP-recommended regimen]{Comparison of six spatially-adjusted RSV prophylaxis regimens to the AAP-recommended regimen, showing statewide superiority and regional non-inferiority of spatially-adjusted regimens. Preventable fraction: the fraction of all RSV-associated hospitalizations occurring while the prophylaxis regimen provides protection against RSV}
\label{fig:coverageByRegionRegimen}
\end{center}
\end{figure*}

@@

The cause of this is in temporal misalignment between the RSV season (delineated by median \Sexpr{seasonThreshold * 100}\%  onset and \Sexpr{(1 - seasonThreshold) * 100}\% offset in each region), and the timing of the AAP-recommended prophylaxis regimen. (Figure \ref{fig:coverageByRegionAAPInsight}.) Fairfield county, owing to its early season onset, sees disproportionately many cases before prophylaxis administration begins; on the other hand, the low-RSV counties, owing to their late season onset and relatively shorter RSV seasons, see almost all of their RSV cases within the prophylaxis window.

\paragraph{Spatially-adjusted regimens are superior to the AAP recommendations}

@@

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=epiyear), width=.3) + 
        geom_point(aes(y=median, x=epiyear), size=.75) +
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_continuous(breaks=seq(1990, 2020, by=5)) +
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks))) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        labs(y=NULL) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank()
        )

@@

Our linear regression of median RSV season onset and offset over three-year periods indicates that season onset has slowly been drifting earlier, at a rate of \Sexpr{formatDrift(onsetByYear)} (\Sexpr{formatDriftCI(onsetByYear, level)}). (Figure \ref{fig:onsetOffsetByYear}) It is unclear whether this is a part of a larger pattern, such as a cycle spanning decades.

@@

Although we were able to reach some conclusions about the impact of calendrical rounding on spatially-adjusted prophylaxis regimens, ultimately our analysis is hamstrung by weekly aggregation of hospital admission data in our initial data. To properly propose and evaluate any fine-tuning to RSV prophylaxis guidelines, we recommend use of daily data.

@@

@@

We have also shown a misalignment between the prophylaxis regimen recommended by the AAP and the timing of the RSV season throughout Connecticut, with most counties' RSV season starting before protection from the first dose of prophylaxis takes hold, and ending before protection from the last dose wanes. This misalignment is particularly notable in Fairfield County, due to its notably early season onset.

@@

However, increasing complexity of practice guidelines carries with it increased cost of implementation, such as training of clinicians and other healthcare workers. It is therefore not apparent from our research alone that the potential benefits would outweigh the costs.

Finally, although we found a weak statewide trend in RSV season onset becoming earlier over time, further analysis did not yield any improvements to prophylaxis based on that trend. Since temporal variation in clinical guidelines would lead to a higher implementation burden than spatial variation, we see no reason to base RSV immunoprophylaxis clinical practice guidelines on temporal variation in RSV reason timing.

>>>>

% TODO: Plot regression line in onset/offset analysis
% TODO clarify all years analysis in figure captions
% TODO Put 1-2% punchline in the abstract, maybe

@@

\abstractsection{Background}Respiratory syncytial virus (RSV) causes seasonal respiratory infection with potentially serious complications in children, and leads to hospitalization rates as high as 50\% in high-risk infants. Safe and effective immunoprophylaxis (palivizumab) is available, but costly. The American Academy of Pediatrics (AAP) recommends palivizumab only for infants at high risk of complications, and only during the RSV season. Although the current AAP guidelines acknowledge the existence of spatial and temporal variation in RSV incidence, they do not recommend spatial or temporal adjustments to immunoprophylaxis regimen outside of Florida.

% TODO: connect better to next paragraph (In order to determine if CT is also variable)

@@

\abstractsection{Results}Using preventable fraction to assess different immunoprophylaxis regimens, we found that (at \Sexpr{level * 100}\% confidence level), regimens adjusted for county-level variation in RSV incidence are superior to the AAP-recommended regimen. When we rounded the timing of those regimens to pragmatic calendrical boundaries (weekly, biweekly, and monthly), their gains over the AAP regimen diminished.

\abstractpara{}Our best-performing spatially adjusted pragmatic alternate to the AAP guidelines was the one based on regional RSV season midpoint with biweekly rounding. In comparison to the AAP recommendation, whose preventable fraction is \Sexpr{regionStrategyCoverage("all", "aap", 0)} (\Sexpr{regionStrategyCoverageCI("all", "aap", 0, level)}), our best alternate yields improvement to \Sexpr{regionStrategyCoverage("all", "countyMiddle", 2)} (\Sexpr{regionStrategyCoverageCI("all", "countyMiddle", 2, level)}) preventable fraction.

\abstractpara{}We also found that alternate regimens adjusted for temporal variation in incidence at the annual level have non-superior preventable fraction to our spatially adjusted regimens.

@@

RSV seasonality has been established throughout the United States, based on National Respiratory and Enteric Virus Surveillance System data. In all states except for Florida, it shows a stable season onset between mid-November and early January and a stable season offset between mid-March and late April.\cite{CentersforDiseaseControlandPreventionCDC:2011wt,CentersforDiseaseControlandPreventionCDC:2013wv}

@@

Motivated by the AAP's recognition of the significance of spatial variation in formulating prophylaxis guidelines, and the known spatial variability of RSV incidence in Connecticut, we aim to determine whether spatial and temporal variability can be used to optimize RSV prophylaxis in Connecticut by asking these questions:

@@

\begin{itemize}
    \item One region for the entire state
    \item One region for each of the counties with high RSV case counts: Hartford County, Fairfield County, and New Haven County. (Figure \ref{fig:map}.)
    \item One discontiguous region comprised of all remaining counties in Connecticut: \Sexpr{gsub("(.*), (.*)", "\\1, and \\2", paste(lowIncidenceCounties, collapse=", "))} (``low-RSV counties'').
\end{itemize}

@@

%TODO citation needed

@@

\subsection{Evaluating spatially adjusted prophylaxis regimens}

@@

In each variant, we again estimated the preventable fraction of all six spatially adjusted regimens.

@@

We then compared those regimens to the one recommended by AAP and to the six spatially adjusted alternates (with no calendrical rounding) discussed above.

@@

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=county), width=.1) + 
        geom_point(aes(y=median, x=county), size=.75) + 
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        labs(aes(x=variable)) +
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_discrete(labels=countyLabels) + 
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks)), sec.axis=sec_axis(trans=~., breaks=seq(1, 52))) + 
        labs(y=NULL) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=30, hjust=1)
        )

@@

    ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_ribbon(aes(ymin=rsv.cum.fit.lower, ymax=rsv.cum.fit.upper, x=time), fill=grey(.25)) +
    	geom_segment(aes(x=start, y=lineYmin, xend=start, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=end, y=lineYmin, xend=end, yend=lineYmax, color=variable), data=spans, size=.375) + 
    	geom_rect(aes(xmin=start, ymin=barYmin, xmax=end, ymax=barYmax, fill=variable, color=variable), data=spans, size=.375) + 
    	geom_segment(aes(x=-Inf, y=unprotected.start, xend=ppx.start, yend=unprotected.start), linetype="11", data=thresholds, size=.375) + 
        geom_segment(aes(x=-Inf, y=unprotected.end, xend=ppx.end, yend=unprotected.end), linetype="11", data=thresholds, size=.375) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.end), data=thresholds, shape=18) + 
        geom_point(aes(x=min(monthBoundaries$min), y=unprotected.start), data=thresholds, shape=18) + 
        scale_y_continuous(labels=latexPercent, breaks=seq(0, 1, by=.2), limits=c(-0.1, 1.1), expand=c(0, 0)) +
        scale_x_continuous(breaks=monthBoundaries$mid, labels=epiWeekLabels, limits=range(c(monthBoundaries$min, monthBoundaries$max)), expand=c(0, 0)) + 
        scale_color_manual(labels=spanLabels, values=twoToneAlt) +
        scale_fill_manual(labels=spanLabels, values=twoToneAlt) +
        guides(color=guide_legend(keywidth=3, keyheight=.4)) + 
        facet_wrap(~county, ncol=5, labeller=labeller(county=countyLabels)) +
        labs(x=NULL, y="Cumulative indicence fraction") + 
        theme(
            panel.grid.major.x=element_blank(),
            legend.title=element_blank(),
            axis.title.y=element_blank(),
            axis.ticks.x=element_blank(),
            axis.text.x=element_text(angle=90, hjust=0.5, vjust=0.5),
            legend.position="bottom"
        )

@@

\setcounter{figure}{4}
\begin{figure*}
\begin{center}
\input{\figuresDir/coverageByRegionRegimen.tex}
\caption[Comparison of six spatially adjusted RSV prophylaxis regimens to the AAP-recommended regimen]{Comparison of six spatially adjusted RSV prophylaxis regimens to the AAP-recommended regimen, showing statewide superiority and regional non-inferiority of spatially adjusted regimens. Preventable fraction: the fraction of all RSV-associated hospitalizations occurring while the prophylaxis regimen provides protection}
\label{fig:coverageByRegionRegimen}
\end{center}
\end{figure*}

@@

This is caused by temporal misalignment between the RSV season (delineated by median \Sexpr{seasonThreshold * 100}\%  onset and \Sexpr{(1 - seasonThreshold) * 100}\% offset in each region), and the timing of the AAP-recommended prophylaxis regimen. (Figure \ref{fig:coverageByRegionAAPInsight}.) Fairfield county, owing to its early season onset, sees disproportionately many cases before prophylaxis administration begins; on the other hand, the low-RSV counties, owing to their late season onset and relatively shorter RSV seasons, see almost all of their RSV cases within the prophylaxis window.

\paragraph{Spatially adjusted regimens are superior to the AAP recommendations}

@@

    plot = ggplot(data) +
        theme_light(base_size=plotTextBaseSize) +
        geom_errorbar(aes(ymin=lower, ymax=upper, x=epiyear), width=.3) + 
        geom_point(aes(y=median, x=epiyear), size=.75) +
        geom_blank(data=limits, aes(y=min)) + 
        geom_blank(data=limits, aes(y=max)) + 
        facet_wrap(~variable, ncol=1, scales="free_y", labeller=onsetOffsetLabeller) +
        scale_x_continuous(breaks=seq(1990, 2020, by=5)) +
        scale_y_continuous(minor_breaks=seq(1, 52), breaks=epiWeekBreaks, labels=rep("\\qquad\\qquad", length(epiWeekBreaks)), sec.axis=sec_axis(trans=~., breaks=seq(1, 52))) + 
        annotation_custom(monthGrob(4), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[4], ymax=monthBoundaries$mid[4]) + 
        annotation_custom(monthGrob(5), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[5], ymax=monthBoundaries$mid[5]) + 
        annotation_custom(monthGrob(9), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[9], ymax=monthBoundaries$mid[9]) + 
        annotation_custom(monthGrob(10), xmin=monthBarX, xmax=monthBarX, ymin=monthBoundaries$mid[10], ymax=monthBoundaries$mid[10]) + 
        labs(y=NULL) + 
        theme(
            panel.grid.major.y=element_blank(),
            panel.grid.minor.y=element_line(size=rel(2)),
            axis.title.x=element_blank()
        )

@@

Our linear regression of median RSV season onset and offset over three-year periods indicates that season onset has slowly been drifting earlier, at a rate of \Sexpr{formatDrift(onsetByYear)} (\Sexpr{formatDriftCI(onsetByYear, level)}). (Figure \ref{fig:onsetOffsetByYear}.) It is unclear whether this is a part of a larger pattern, such as a cycle spanning decades.

@@

Although we were able to reach some conclusions about the impact of calendrical rounding on spatially adjusted prophylaxis regimens, ultimately our analysis is hamstrung by weekly aggregation of hospital admission data in our initial data. To properly propose and evaluate any fine-tuning to RSV prophylaxis guidelines, we recommend use of daily data.

@@

% TODO Revisit

@@

Furthermore, we have shown a misalignment between the prophylaxis schedule recommended by the AAP and the timing of the RSV season throughout Connecticut, with most counties' RSV season starting before protection from the first dose of prophylaxis takes hold, and ending before protection from the last dose wanes. This misalignment is particularly notable in Fairfield County, due to its early season onset.

@@

Increasing complexity of practice guidelines carries with it increased cost of implementation, such as training of clinicians and other healthcare workers. It is therefore not apparent from our research alone that the potential benefits would outweigh the costs.

Finally, although we found a weak statewide trend in RSV season onset becoming earlier over time, further analysis did not yield any improvements to prophylaxis based on that trend. Since temporal variation in clinical guidelines would lead to a higher implementation burden than spatial variation, we see no reason to base RSV immunoprophylaxis clinical practice guidelines on temporal variation in RSV season timing.
